---
title: "Beta Distribution"
author: "Aaron Mulvey"
date: "9/27/2022"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)

```

## R Markdown

The Beta distribution Beta(a, b) has the following Probability Density Function (PDF):
 $$f (x) = cx^{\alpha−1}(1 −x)^{\beta−1} $$
 $$ 0 < x < 1,$$
with parameters a > 0 and b > 0, and a constant c chosen to make it a valid PDF. Let X ∼Beta(a, b) with
parameters a = 6 and b = 2.

1. Find c for the given parameters
    
This is the beta fuction:

$$f (x; \alpha ;\beta) = cx^{\alpha−1}(1 −x)^{\beta−1} $$
c makes the function valid so the integral in the support must be = 1 so c is equal to:
$$c \int_{0}^{1} f (x; \alpha ;\beta) dx =1 \implies c = \frac{1}{\int_{0}^{1} f (x; \alpha ;\beta)dx}$$
so that means that:

$$c = \frac{1}{B(\alpha, \beta)}$$
I substitute in my formula and I get the following:

$$ f (x; \alpha ;\beta) = \frac{1}{B(\alpha, \beta)}x^{\alpha−1}(1 −x)^{\beta−1}$$
I saw that I can also do it using the gamma distribution, looks like this:

$$ f (x; \alpha ;\beta)= \frac{x^{\alpha−1}(1 −x)^{\beta−1}}{\int_{0}^{1} \omega^{\alpha−1}(1 −\omega)^{\beta−1}d\omega} $$
$$ = \frac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)} x^{\alpha−1}(1 −x)^{\beta−1}$$

```{r}
a <- 6
b <- 2

constant_with_BETA <- function(a, b){
  constant <-  1/beta(a,b)
  return (constant)
}

constant_with_GAMMA <- function(a, b){
  constant <-  gamma(a + b)/(gamma(a) * gamma(b))
  return (constant)
}
constant_with_GAMMA(a,b)
constant_with_BETA(a,b)

```

2. Find P (0.2 < X < 0.8)

$$P(\alpha< X < \beta) = P(\alpha < X \le \beta) = P(\alpha \le X < \beta) = P(\alpha \le X \le \beta)$$
This is because:

$$P(X = \alpha) = 0$$
Because:

$$\int_{\alpha}^{\alpha} f(x)dx = 0$$
Also using the naive definition of probability:

$$P(X = \alpha) = \frac{\alpha}{\Omega}$$
Since we are working with continous random variables the sample set is infinite so we have that:

$$P(X = \alpha) = \frac{\alpha}{\infty} = 0$$



```{r}
integrate(dbeta, lower=0.2, upper=0.8, a, b)
```

3.1 Find E(X)

formula of expectation:
$$ \mu = E[X] = \int_{0}^{1} xf (x; \alpha ;\beta) dx$$
I expand it.
$$  = \int_{0}^{1} x \frac{x^{\alpha−1}(1 −x)^{\beta−1}}{B(\alpha, \beta)}dx$$
And I get the following:
$$ = \frac{\alpha}{\alpha + \beta} $$
Take comun divisor alpha:
$$ = \frac{1}{1 + \frac{\beta}{\alpha}}$$
```{r}

Expected_value_X <- function(a,b){
  return (1/(1 + (b/a)))
}

Expected_value_X(a,b)

```

3.2 Find Var(X)

$$var(X) =E[(X - \mu)^2] = \frac{\alpha \beta}{(\alpha + \beta)^2(\alpha + \beta + 1)}$$

```{r}
var_X <- function(a, b){
  return ((a * b) / ((a + b)^2 * (a + b + 1)))
}

var_X(a,b)
```

Let X be a random variable with PDF f , an example PDF is depicted in Figure 1. Here is what you know:

  You can evaluate f for specific values x
  You do not know the Cumulative Distribution Function (CDF) of f
  You have a random number generator that can draw from the uniform distribution U Unif(a, b),     for any a and b
  You can find the absolute maximum of f
  The support of f is between 0 and 10
  
(a) Create your own version of Figure 1 (it doesn’t have to be exactly the same function). Implement (in code) a simulation experiment to compute the following probability for your own PDF f, P(3 ≤ X < 6).

First, I will create my function:

$$f(x) = \frac{\sqrt{x}}{21.08185}$$

```{r}
suport <- 0
SUPORT <- 10
a <- 3
b <- 6

fun <- function(x) ifelse((x > 10),0, (sqrt(x)/21.08185))

curve(fun, 0, 15)

```
Then I will take the maximun point in the interval given, I did this because I can calculate the area of the rectangle knowing the interval (a, b) and the maximun height of my function.

```{r}

FIND_MAXIMUN <- function(fun, a, b, n){
  maximun = fun(a)
  secuencia <- seq(a, b, by = n)
  for (i in secuencia) {
      maximun <- ifelse(fun(i) > maximun, maximun <- fun(i), maximun <- maximun)
  }
  return (maximun)
}

curve(fun, 0, 15)
abline(h = 0, lty = 2)
abline(v = c(3, 6), lty = 2)
abline(h = FIND_MAXIMUN(fun, a , b, 0.01), lty = 2)

```
What I am going to do now is to throw random points in that square and I will store the poits that are below my function whith that I will calculate the ratio and I will multiply the area of the rectangle with that ratio so that I will get the integral in the given point.

```{r}

RANDOM_POINTS <- function(fun, a, b, n, visualice){
 
  
  x = c()
  Fx = c()
  color = c()
  true <- 0
  false <- 0
  
  maximun <- FIND_MAXIMUN(fun,a,b,0.01)
  
  for (i in 1:n) {

  x[i] = runif(1, a, b)
  Fx[i] = runif(1, 0, maximun)
  color[i] = ifelse(Fx[i] > fun(x[i]), TRUE, FALSE)
  ifelse(color[i] == T, true <- true + 1, false <-  false + 1)
  
  }

  
  if(visualice == T){
    plot(x, Fx, col = ifelse(color == TRUE, 'red', 'green'), pch = 19, cex = 0.2) 
  }else{
    return (((b-a) * maximun) * (false/n))
  }
  
  
}

RANDOM_POINTS(fun, a, b, 100000, T)

```

Now I will check if the result of the integral is correct, whith the method that R provide us.

```{r}

integrate(fun, lower = a, upper = b)
RANDOM_POINTS(fun, a, b, 10000, F)


```


(b) What components of your simulation have impact on approximation accuracy?

I will do an experiment increasing the number of points that I input to my simulation, whit this I will plot in a graph for see how the number of points affects my approximation, for calculating the error I will use the value that the native function of R provide us.


```{r}

n <- 10
N <- 1000
delta <- 10
error <- c()

EXPERIMENT <- function(n, N, delta, integral, fun, a, b){
  
  increment <- seq(n, N, by = delta)
    for (i in 1:length(increment)) {
      aproximation <- RANDOM_POINTS(fun, a, b, increment[i], F)
      error[i] <- (integral -  aproximation)
    }
  plot(increment, error) 
  abline(h = 0, lty = 2)
  return (data.frame(error, increment))
}

integral <- integrate(fun, lower = a, upper = b)

integral <- integral$value
  
error <- EXPERIMENT(n, N, delta, integral, fun, a , b)


```

As we can see as the number of points increments the error goes decreasing, so the number of points is the component that affects my simulation acuracy.

Same thing happends when I use the squared error:

```{r}

error <- error %>%
  mutate(error_squared = error^2)

plot(x = error$increment, y = error$error_squared)
abline(h = 0, lty = 2)
```

I have not mentioned but the PDF of my function is valid because:
```{r}

integrate(fun, lower = suport, upper = SUPORT)

```
And also is bigger than 0
